---
title: "sales forecasting  using multiple linear regression"
author: "JOSEPH MUTUA"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction

The objective of this study is to analyze factors that influence weekly retail store sales and to develop a multiple linear regression model for predicting sales performance.

#Dataset Description

The dataset contains weekly observations for 45 stores with the following variables:

- Weekly_Sales (target variable)
- Store
- Holiday_Flag
- Temperature
- Fuel_Price
- CPI
- Unemployment

# Research Question

Do economic and environmental factors significantly affect weekly sales?

#Statistical Model

We use Multiple Linear Regression:

Weekly_Sales = β0 + β1Store + β2Holiday + β3Temperature + β4Fuel_Price + β5CPI + β6Unemployment + ε

## Hypotheses

H0: All regression coefficients = 0  
H1: At least one coefficient ≠ 0

# Data Loading
```{r}
sales <- read.csv("sales.csv")
sales$Store <- as.factor(sales$Store)
sales$Holiday_Flag <- as.factor(sales$Holiday_Flag)
sales$Date <- as.Date(sales$Date)

```



#Exploratory Analysis
```{r}
summary(sales)
```



# Model Fitting
```{r}
model <- lm(Weekly_Sales ~ Store + Holiday_Flag + Temperature +
              Fuel_Price + CPI + Unemployment, data = sales)
summary(model)
```



#Model Evaluation
```{r}
pred <- predict(model, sales)
rmse <- sqrt(mean((sales$Weekly_Sales - pred)^2))
rmse
```



# Results and Interpretation
The overall regression model was statistically significant (p < 0.001), indicating that the predictors jointly explain variations in weekly sales.

Key findings:

Holiday periods increase sales

Higher fuel prices reduce sales

Higher unemployment reduces sales

Temperature has a negative relationship with sales

R² = 0.92 indicating strong explanatory power

# Conclusion
The analysis demonstrates that weekly sales are significantly influenced by store characteristics and economic conditions. The regression model provides reliable predictions and can support business decision-making such as inventory planning and staffing during holidays.

